---
title: Weather-Analysis
author: ~
date: '2017-06-20'
slug: weather-analysis
categories: []
tags: []
Categories: []
Description: ''
Tags: [weather]
---
For this analysis I scraped weather data for the year 2016 from WeatherUndergrounds historical records. 
https://www.wunderground.com/history/airport/FACT/2016/01/01/DailyHistory.html?format=1 

After compiling all of the 2016 data into a CSV, I edited the data by removing any entries that held an "NA" value. For Numeric records it appears that if data was not be obtained for some entry, Weather Underground recorded as having a value of -9999. The big negative values could impact the accuracy of my models, so I removed that as well.

For my model, I decided to use a GAM. So, I did some graphing to see which catgories seem to have some sort of significance.
```{r cond plot, echo = FALSE}
library(ggplot2)
weather_data = read.csv("C:/Users/Student/Documents/R/Data Analysis/weather2016edited.csv")
ggplot2::ggplot(weather_data,ggplot2::aes(x=TemperatureC, fill = Conditions)) +ggplot2::geom_density(alpha=.5) + ggplot2::theme_classic()
```
While the conditions density graph was not particularly enlightening, it does seem to suggest that certain weather conditions will indicate certain temperature ranges. 
```{r fit_cond, echo=TRUE}
#library(gam)
#fit = gam(TemperatureC ~ Conditions, data = weather_data)
```

```{r temperature, echo = FALSE, warning=FALSE}
weather_data$month = factor(weather_data$month, unique(weather_data$month))
ggplot2::ggplot(weather_data,ggplot2:: aes(x = TimeSAST, y = TemperatureC, color = month)) + ggplot2:: geom_point(alpha = .5) + ggplot2:: geom_smooth() + ggplot2:: theme_classic()
```

The Above graph, however, says a great deal. There is a very obvious relationship between Time and Temperature as well as Month and temperature, with the temperature rising during daylight hours and falling during the night, as well as the temperatures being higher in summer months and lower in winter months. Considering the relatively visible trend, I need to add them into my model. However, because they are quantitative rather than quantitative, I need to determine how many degrees of freedom to give their functions. Also given the curved nature of the relationship, it is likely that theses functions will have relatively high degrees of freedom. First I checked the Time fit function.
```{r Time and Month, echo=FALSE}
library(gam)
weather_data$month = as.numeric(weather_data$month)
fit_temp0 = gam(TemperatureC ~ Conditions, data = weather_data)
fit_temp1 = gam(TemperatureC ~ Conditions + poly(TimeSAST,1), data = weather_data)
fit_temp2 = gam(TemperatureC ~ Conditions + poly(TimeSAST,2), data = weather_data)
fit_temp3 = gam(TemperatureC ~ Conditions + poly(TimeSAST,3), data = weather_data)
fit_temp4 = gam(TemperatureC ~ Conditions + poly(TimeSAST,4), data = weather_data)
fit_temp5 = gam(TemperatureC ~ Conditions + poly(TimeSAST,5), data = weather_data)
fit_temp6 = gam(TemperatureC ~ Conditions + poly(TimeSAST,6), data = weather_data)
fit_temp7 = gam(TemperatureC ~ Conditions + poly(TimeSAST,7), data = weather_data)
fit_temp8 = gam(TemperatureC ~ Conditions + poly(TimeSAST,8), data = weather_data)
fit_temp9 = gam(TemperatureC ~ Conditions + poly(TimeSAST,9), data = weather_data)
fit_temp10 = gam(TemperatureC ~ Conditions + poly(TimeSAST,10), data = weather_data)
fit_temp11 = gam(TemperatureC ~ Conditions + poly(TimeSAST,11), data = weather_data)
fit_temp12 = gam(TemperatureC ~ Conditions + poly(TimeSAST,12), data = weather_data)

anova(fit_temp0,fit_temp1,fit_temp2,fit_temp3,fit_temp4,fit_temp5,fit_temp6,fit_temp7,fit_temp8,fit_temp9,fit_temp10,fit_temp11,fit_temp12)

```

Surprisingly, for the Time fit function The model continued to be statistically relevant up until an eleventh degree polynomial. Then I looked into the polynomial month function

```{r month, echo=FALSE}
fitm0 = gam(TemperatureC ~ Conditions + poly(TimeSAST,11) + poly(Humidity,4), data = weather_data)
fitm1 = gam(TemperatureC ~ Conditions + poly(TimeSAST,11) + poly(Humidity,4) + poly(month,1), data = weather_data)
fitm2 = gam(TemperatureC ~ Conditions + poly(TimeSAST,11) + poly(Humidity,4) + poly(month,2), data = weather_data)
fitm3 = gam(TemperatureC ~ Conditions + poly(TimeSAST,11) + poly(Humidity,4) + poly(month,3), data = weather_data)
fitm4 = gam(TemperatureC ~ Conditions + poly(TimeSAST,11) + poly(Humidity,4) + poly(month,4), data = weather_data)
fitm5 = gam(TemperatureC ~ Conditions + poly(TimeSAST,11) + poly(Humidity,4) + poly(month,5), data = weather_data)

anova(fitm0,fitm1,fitm2,fitm3,fitm4,fitm5)

```

From this analysis it appears that the month function only requires a degree four polynomial. From here I looked into the relationship between humidity and temperature.

```{r humid plot, echo =FALSE}
library(RColorBrewer)
ggplot2::ggplot(weather_data, ggplot2::aes(x=Humidity,y=TemperatureC, color = TemperatureC)) + ggplot2:: geom_point() + geom_smooth() + ggplot2:: theme_classic()
```

This graph definitely presents a pretty clear and relatively linear realtionship between the two variables, with temperature falling as Humidity rose. Thus I explored the fit function for Humidity.

```{r humid anova, echo = FALSE}
fit_temp1 = gam(TemperatureC ~ Conditions + poly(TimeSAST,11), data = weather_data)
fit_h_1 = gam(TemperatureC ~ Conditions + poly(TimeSAST,11) + poly(Humidity,1), data = weather_data)
fit_h_2 = gam(TemperatureC ~ Conditions + poly(TimeSAST,11) + poly(Humidity,2), data = weather_data)
fit_h_3 = gam(TemperatureC ~ Conditions + poly(TimeSAST,11) + poly(Humidity,3), data = weather_data)
fit_h_4 = gam(TemperatureC ~ Conditions + poly(TimeSAST,11) + poly(Humidity,4), data = weather_data)
fit_h_5 = gam(TemperatureC ~ Conditions + poly(TimeSAST,11) + poly(Humidity,5), data = weather_data)
anova(fit_temp1,fit_h_1,fit_h_2,fit_h_3,fit_h_4,fit_h_5)
```

Here we see that the humidity model only requires one degree, which makes sense considering how linear the Temp Vs. Humidity graph appears to be. Finally, I looked into the relationship between atmospheric pressure and temperature.

```{r pressure}
ggplot2::ggplot(weather_data, ggplot2::aes(x=Pressure, y = TemperatureC, color = TemperatureC)) + ggplot2::geom_point() +ggplot2::theme_classic() 
```

From this graph it does seem like there is some kind of relationship here, though it is rather unclear, exactly what it is.

```{r pressure stat, echo=FALSE}
pres0 = gam(TemperatureC ~ Conditions + poly(TimeSAST,11) + poly(Humidity,4) + poly(month,4), data = weather_data)
pres1 = gam(TemperatureC ~ poly(Pressure,1) + Conditions + poly(TimeSAST,11) + poly(Humidity,4) + poly(month,4), data = weather_data)
pres2 = gam(TemperatureC ~ poly(Pressure,2) + Conditions + poly(TimeSAST,11) + poly(Humidity,4) + poly(month,4), data = weather_data)
pres3 = gam(TemperatureC ~ poly(Pressure,3) + Conditions + poly(TimeSAST,11) + poly(Humidity,4) + poly(month,4), data = weather_data)
pres4 = gam(TemperatureC ~ poly(Pressure,4) + Conditions + poly(TimeSAST,11) + poly(Humidity,4) + poly(month,4), data = weather_data)
pres5 = gam(TemperatureC ~ poly(Pressure,5) + Conditions + poly(TimeSAST,11) + poly(Humidity,4) + poly(month,4), data = weather_data)


anova(pres0,pres1,pres2,pres3,pres4,pres5)
```

From this analysis, it appears that the function is statistically significant up to four degrees of freedom.

After exploring all of the different potential factors in my model, I built my model.

```{r predictions, echo =TRUE}
fit.weather = gam :: gam(TemperatureC ~ poly(Pressure,3) + Conditions + poly(TimeSAST,11) + poly(Humidity,1) + poly(month,4), data = weather_data)
```

After building this GAM model, I ran the model 20 times and found the average error for each individual trial and then determiend the average error for the model as a whole

```{r error, echo=FALSE}
seeds = sample(1:10000,20,replace = TRUE)
error = c()
for(i in seeds){
  weather_data = read.csv("C:/Users/Student/Documents/R/Data Analysis/weather2016edited.csv")
  set.seed(i)
  weather_data$month = as.numeric(weather_data$month)
  train.index = sample(c(T, F), nrow(weather_data), replace = TRUE, prob = c(0.8, 0.2))
  weather_data = split(weather_data, train.index)
  names(weather_data) = c("test", "train")
  remove(train.index)
  fit.weather = gam :: gam(TemperatureC ~ poly(Pressure,3) + Conditions + poly(TimeSAST,11) + poly(Humidity,1) + poly(month,4), data = weather_data$train)
  weather_data$test$predict = predict(fit.weather, newdata = weather_data$test)
  weather_data$test$error = abs(weather_data$test$predict - weather_data$test$TemperatureC)/weather_data$test$TemperatureC
  error = c(error,mean(weather_data$test$error))
}
sprintf("the average error for this model with 10 repetitions is %f, the maximum error is %f, and the minimum error is %f", mean(error), max(error), min(error))
```

The average error for this model appears to be around 12%, which, considering the chaotic nature of weather, I beleive to be a pretty resonable margin. From this, I think it can be concluded that a GAM model is rather effective in predicting temperatures within a given year. However, it remains to be seen whether or not this model will be effective in extrapolating temperature data into the future.



